---
title: "Regression Model Project"
author: "tuhin mahmud"
date: "Thursday, November 20, 2014"
output: html_document
---
#Synopsis
The purpose of this project is to look at a data set of a collection of cars.We are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). We want to answer following two questions:

1. "Is an automatic or manual transmission better for MPG"
1. "Quantify the MPG difference between automatic and manual transmissions"
##Summary
Studying ther results of the linear regression and multi variable linear regression we find that automatic transmission is better for MPG but impact of transmission type on MPG is lowered when considered with other variables.

For linear regression the MPG increasses  7.245 MPG , whereas if we consider other important variables like cylinder type , weight  and horse power the increase in MPG is about 1.809 MPG.

#Data Analysis
```{r}
##clean data
library(datasets)
data<-mtcars
data$cyl<-factor(data$cyl)
data$am<-factor(data$am)
levels(data$am)<-c("automatic","manual")
data$gear<-factor(data$gear)
data$carb<-factor(data$carb)
```
#Variable Selection 
using method suggested in http://www.statmethods.net/stats/regression.html
```{r}
# Stepwise Regression
library(MASS)
fit <- lm(mpg~am+cyl+wt+hp,data=data)
step <- stepAIC(fit, direction="both")
step$anova # display results


```
# Multvariable Regression Analysis
```{r}
#sm = step(lm(data = mtcars, mpg ~ .),trace=0,steps=10000)
sm = step(lm(data = mtcars, mpg ~ .))
summary(sm)
```
# Build Linear Regression Models
```{r}

model1<-lm(mpg~am,data=data)
summary(model1)$coef

model2<-lm(mpg~cyl,data=data)
summary(model2)$coef

model3<-lm(mpg~cyl+wt,data=data)
summary(model3)$coef

model4<-lm(mpg~cyl+wt+hp,data=data)
summary(model4)$coef

model5<-lm(mpg~am + cyl + wt + hp,data=data)
summary(model5)$coef

model6<-lm(mpg~.,data=data)
summary(model6)$coef

model7<-lm(mpg~ wt + qsec + am,data=data)
summary(model7)$coef

#Pair wise anova comparison of the models
anova(model1,model2)
anova(model1,model3)
anova(model1,model4)
anova(model1,model5)
anova(model1,model6)
anova(model1,model7)

```
## Plot Data
```{r}
library(ggplot2)
qplot(x=wt,y=mpg,data=data,colour=am,facets=.~cyl, main="MPG  vs wt graph")

#ref: http://www.statmethods.net/graphs/scatterplot.html
# Simple Scatterplot
attach(mtcars)
plot(wt, mpg, main="Scatterplot Example", 
     xlab="Car Weight ", ylab="Miles Per Gallon ", pch=19)
#simple scatterplot click to view
# Add fit lines
abline(lm(mpg~wt), col="red") # regression line (y~x) 
lines(lowess(wt,mpg), col="blue") # lowess line (x,y)

boxplot(mpg~am, data = mtcars,
        col = c("dark grey", "light grey"),
        xlab = "Transmission",
        ylab = "Miles per Gallon",
        main = "MPG by Transmission Type")
```


#Analyze Different Models
Analyze the five models with plot 
```{r}
#model1
summary(model1)
#layout(matrix(c(1,2,3,4),2,2)) 
#plot(model1)
#model2
summary(model2)
#layout(matrix(c(1,2,3,4),2,2)) 
#plot(model2)

#model3
summary(model3)
#layout(matrix(c(1,2,3,4),2,2)) 
#plot(model3)

#model4
summary(model4)
#layout(matrix(c(1,2,3,4),2,2)) 
#plot(model4)
#model5
summary(model5)
#layout(matrix(c(1,2,3,4),2,2)) 
#plot(model5)

#model6
summary(model6)
#layout(matrix(c(1,2,3,4),2,2)) 
#plot(model6)

# http://people.duke.edu/~rnau/Notes_on_linear_regression_analysis--Robert_Nau.pdf
# Comparing regression model using R squared metric
adj.r2<-c(summary(model1)$adj.r.squared,
          summary(model2)$adj.r.squared,
          summary(model3)$adj.r.squared,
          summary(model4)$adj.r.squared,
          summary(model5)$adj.r.squared,
          summary(model6)$adj.r.squared,
          summary(model7)$adj.r.squared
          )
names(adj.r2)<-c("model1","model2","model3","model4",
                 "model5","model6","model7")
print(adj.r2)
```
#Results
From the Adjusted R^2 results and study of the anova and also as suggested by Stepwise Regression study Model5 turns out to be the best fit.

Model has Adjusted R-squared value of  0.84 with p-value 1.51e-10, so we will need to reject the null hypothesis and claim that our multivalue model is significantly different from the simple linear regression model.

From the above best fit model and the Esitmate coefficient of "ammanual" is 
1.8092, which is the quantity of impact of transmission type on MPG.

Similar coefficient of the simple linear regression model1 shows the impact of the transmission type is 7.245.

```{r}
summary(model5)
layout(matrix(c(1,2,3,4),2,2)) 
plot(model5)

```
